---
title: "publicschools"
output:
  html_document: default
  pdf_document: default
date: '2022-05-06'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## best subset selection

```{r chunk1}
library(ISLR2)
public <- read.csv("updatedpublicschools.csv")
attach(public)
public <- na.omit(public)
```

```{r chunk2}
lm.fit = lm(X..Graduated ~ X..Economically.Disadvantaged + X..English.Language.Learner + X..Students.With.Disabilities + X..High.Needs  + X..Hispanic + X..White + X..Males)
summary(lm.fit)

```
- The residuals plot shows some curvature, thus a linear model may not be reasonable for this data. The data does appear to be heteroscedastic as indicated by the fanning of the residuals.
- The Q-Q plot indicates that normality of the `X..Graduated` data may be not be a reasonable assumption. 
- - The standardized residuals versus leverage plot indicate that some of the observations appear to be outliers or high leverage points.


```{r chunk3}
library(leaps)
regfit.full <- regsubsets(X..Graduated ~ X..Economically.Disadvantaged + X..First.Language.Not.English + X..English.Language.Learner + X..Students.With.Disabilities + X..High.Needs + X..African.American + X..Asian + X..Hispanic + X..White + X..Native.American + X..Native.Hawaiian..Pacific.Islander + X..Multi.Race..Non.Hispanic + X..Males + X..Females, public)
summary(regfit.full)

```

```{r chunk4}
regfit.full <- regsubsets(X..Graduated ~ X..Economically.Disadvantaged + X..First.Language.Not.English + X..English.Language.Learner + X..Students.With.Disabilities + X..High.Needs + X..African.American + X..Asian + X..Hispanic + X..White + X..Native.American + X..Native.Hawaiian..Pacific.Islander + X..Multi.Race..Non.Hispanic + X..Males + X..Females, public, nvmax = 14)

reg.summary <- summary(regfit.full)
reg.summary

```


```{r chunk5}
reg.summary$adjr2
```


```{r chunk6}
par(mfrow = c(2,2))
plot(reg.summary$rss, xlab = "Number of Variables", ylab = "RSS", type = "l")
points(14, reg.summary$rss[14], col = "red", cex = 2, pch = 20)

plot(reg.summary$adjr2, xlab = "Number of Variables", ylab = "Adjusted RSq", type = "l")
points(7, reg.summary$adjr2[7], col = "red", cex = 2, pch = 20)

plot(reg.summary$cp, xlab = "Number of Variables", ylab = "Cp", type = "l")
points(7, reg.summary$cp[7], col = "red", cex = 2, pch = 20)

plot(reg.summary$bic, xlab = "Number of Variables",ylab = "BIC", type = "l")
points(5, reg.summary$bic[5], col = "red", cex = 2, pch = 20)

plot(reg.summary$rsq, xlab = "Number of Variables", ylab = "RSq", type = "l")
points(14, reg.summary$rsq[14], col = "red", cex = 2, pch = 20)

```

```{r chunk7}
which.max(reg.summary$adjr2)
which.min(reg.summary$rss)
which.min(reg.summary$cp)
which.min(reg.summary$bic)
which.max(reg.summary$rsq)
```


```{r chunk8}
plot(regfit.full, scale = "r2")
plot(regfit.full, scale = "adjr2")
plot(regfit.full, scale = "Cp")
plot(regfit.full, scale = "bic")
```

```{r chunk9}
coef(regfit.full, 7)
```

## forward & backward stepwise selection

```{r chunk10}
regfit.fwd <- regsubsets(X..Graduated ~ X..Economically.Disadvantaged + X..First.Language.Not.English + X..English.Language.Learner + X..Students.With.Disabilities + X..High.Needs + X..African.American + X..Asian + X..Hispanic + X..White + X..Native.American + X..Native.Hawaiian..Pacific.Islander + X..Multi.Race..Non.Hispanic + X..Males + X..Females, data = public,
    nvmax = 14, method = "forward")
summary(regfit.fwd)

regfit.bwd <- regsubsets(X..Graduated ~ X..Economically.Disadvantaged + X..First.Language.Not.English + X..English.Language.Learner + X..Students.With.Disabilities + X..High.Needs + X..African.American + X..Asian + X..Hispanic + X..White + X..Native.American + X..Native.Hawaiian..Pacific.Islander + X..Multi.Race..Non.Hispanic + X..Males + X..Females, data = public,
    nvmax = 14, method = "backward")
summary(regfit.bwd)
```

```{r chunk11}
coef(regfit.full, 7)
coef(regfit.fwd, 7)
coef(regfit.bwd, 7)
```

## VALIDATION SET APPROACH 

```{r chunk12}
set.seed(1)
train <- sample(c(TRUE, FALSE), nrow(public), replace = TRUE)
test <- (!train)

regfit.best <- regsubsets(X..Graduated ~ X..Economically.Disadvantaged + X..First.Language.Not.English + X..English.Language.Learner + X..Students.With.Disabilities + X..High.Needs + X..African.American + X..Asian + X..Hispanic + X..White + X..Native.American + X..Native.Hawaiian..Pacific.Islander + X..Multi.Race..Non.Hispanic + X..Males + X..Females, data = public[train, ], nvmax = 14)

test.mat <- model.matrix(X..Graduated ~ X..Economically.Disadvantaged + X..First.Language.Not.English + X..English.Language.Learner + X..Students.With.Disabilities + X..High.Needs + X..African.American + X..Asian + X..Hispanic + X..White + X..Native.American + X..Native.Hawaiian..Pacific.Islander + X..Multi.Race..Non.Hispanic + X..Males + X..Females, data = public[test, ])

val.errors <- rep(NA, 14)
for (i in 1:14) {
 coefi <- coef(regfit.best, id = i)
 pred <- test.mat[, names(coefi)] %*% coefi
 val.errors[i] <- mean((public$X..Graduated[test] - pred)^2)
}
```

```{r chunk13}
val.errors
which.min(val.errors)
coef(regfit.best, 3)
```

```{r chunk14}
 predict.regsubsets <- function(object, newdata, id, ...) {
  form <- as.formula(object$call[[2]])
  mat <- model.matrix(form, newdata)
  coefi <- coef(object, id = id)
  xvars <- names(coefi)
  mat[, xvars] %*% coefi
 }

regfit.best <- regsubsets(X..Graduated ~ X..Economically.Disadvantaged + X..First.Language.Not.English + X..English.Language.Learner + X..Students.With.Disabilities + X..High.Needs + X..African.American + X..Asian + X..Hispanic + X..White + X..Native.American + X..Native.Hawaiian..Pacific.Islander + X..Multi.Race..Non.Hispanic + X..Males + X..Females, data = public, nvmax = 19)
coef(regfit.best, 3)
```
## CROSS VALIDATION 

```{r chunk15}
k <- 10
n <- nrow(public)
set.seed(1)
folds <- sample(rep(1:k, length = n))
cv.errors <- matrix(NA, k, 14,
    dimnames = list(NULL, paste(1:14)))
```


```{r chunk16}
for (j in 1:k) {
  best.fit <- regsubsets(X..Graduated ~ X..Economically.Disadvantaged + X..First.Language.Not.English + X..English.Language.Learner + X..Students.With.Disabilities + X..High.Needs + X..African.American + X..Asian + X..Hispanic + X..White + X..Native.American + X..Native.Hawaiian..Pacific.Islander + X..Multi.Race..Non.Hispanic + X..Males + X..Females,
       data = public[folds != j, ],
       nvmax = 14)
  for (i in 1:14) {
    pred <- predict(best.fit, public[folds == j, ], id = i)
    cv.errors[j, i] <-
         mean((public$X..Graduated[folds == j] - pred)^2)
   }
 }

```


```{r chunk17}
reg.best <- regsubsets(X..Graduated ~ X..Economically.Disadvantaged + X..First.Language.Not.English + X..English.Language.Learner + X..Students.With.Disabilities + X..High.Needs + X..African.American + X..Asian + X..Hispanic + X..White + X..Native.American + X..Native.Hawaiian..Pacific.Islander + X..Multi.Race..Non.Hispanic + X..Males + X..Females, data = public,
    nvmax = 14)
coef(reg.best, 7)
summary(reg.best)
```



